---
title: "Combined files"
author: "Christine Hathaway"
date: "January 4, 2020"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r include = TRUE}
#Set the working directory
setwd("C:/Users/Christine/Documents/Bellevue/DSC 630/Project/Datasets")
```


```{r include = FALSE}
#Install packages needed for file
library(ggplot2)
library(reshape)
library(reshape2)
library("ggpubr")
require(scales)
library("pastecs")
library("Hmisc")
library("caret")
library("plyr")
```

Import data from personsxcsv and sampleadult files

```{r echo = TRUE}
#Import CondPerson.csv file
person2 <- read.csv(file = "C:/Users/Christine/Documents/Bellevue/DSC 630/Project/Datasets/CondPerson.csv", header = TRUE, na.strings = " ")

#Import condsamadult.csv file
adult2 <- read.csv(file = "C:/Users/Christine/Documents/Bellevue/DSC 630/Project/Datasets/condsamadult.csv", header = TRUE, na.strings = " ")
```


Create new file to link person with sampleadult

```{r echo = TRUE }
new <-merge(x = person2, y = adult2, by="ID")

head(new)

describe(new)

```

Create new variables

```{r echo = TRUE }
new$histop1 <- with(new, ifelse(new$HISTOP4 == 1, 1, 0))
new$histop2 <- factor(new$histop1)

new$need1 <- with(new, ifelse(new$PNMED12M == 1, 1, 0))
new$need2 <- factor(new$need1)

new$delay1 <- with(new, ifelse(new$PDMED12M == 1, 1, 0))
new$delay2 <- factor(new$delay1)

new$notcov1 <- with(new, ifelse(new$NOTCOV == 2, 1, 0))
new$notcov2 <- factor(new$notcov1)

new$pdsick1 <- with(new, ifelse(new$PDSICKA == 1, 1, 0))
new$pdsick2 <- factor(new$pdsick1)

new$sex2<- factor(new$SEX.x)
new$sex1 <- new$sex2.as.integer

head(new)
```


Model for have sick pay but delayed care

```{r echo = TRUE }
sickmodel <- glm(delay2 ~ pdsick2 + sex2 + AGE_P, data = new, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(sickmodel)

```



Model  for have sick pay but didn't get needed care

```{r echo = TRUE }
sickmodel2 <- glm(need2 ~ pdsick2 + sex2 + AGE_P, data = new, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(sickmodel2)

```

Create Training and Testing Datasets

```{r echo = TRUE }
#create a list of random number ranging from 1 to number of rows from actual data and 70% of the data into training data  

data1 = sort(sample(nrow(new), nrow(new)*.7))

#creating training data set by selecting the output row values  
trainnew<-new[data1,]

#creating test data set by not selecting the output row values
testnew<-new[-data1,]

```



Model for have sick pay but delayed care using training data

```{r echo = TRUE }
sickmodeltrain <- glm(delay2 ~ pdsick2 + sex2 + AGE_P, data = trainnew, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(sickmodeltrain)

```

Make predictions

```{r echo = TRUE }
#Make predictions
sickpredict <-  predict(sickmodeltrain, testnew, type = "terms")

head(sickpredict)

summary(sickpredict)
```



Model  for have sick pay but didn't get needed care using training data

```{r echo = TRUE }
sickmodel2train <- glm(need2 ~ pdsick2 + sex2 + AGE_P, data = trainnew, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(sickmodel2train)

```

Make predictions

```{r echo = TRUE }
#Make predictions
sick2predict <- predict(sickmodel2train, testnew, type = "terms")

head(sick2predict)

summary(sick2predict)
```


