---
title: "Person"
author: "Christine Hathaway"
date: "December 24, 2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = TRUE}
#Set the working directory
setwd("C:/Users/Christine/Documents/Bellevue/DSC 630/Project/Datasets")
```


```{r include = FALSE}
#Install packages needed for file
library(ggplot2)
library(reshape)
library(reshape2)
library("ggpubr")
require(scales)
library("pastecs")
library("Hmisc")
library("caret")
library("plyr")
```

Import data from personsxcsv file

```{r echo = TRUE }
#Import personsx.csv file
person <- read.csv(file = "C:/Users/Christine/Documents/Bellevue/DSC 630/Project/Datasets/personsx.csv", header = TRUE, na.strings = " ")
```

Display first five records of file

```{r echo = TRUE }
head(person)
```

Run stat.desc() function of file

```{r echo = TRUE }
stat.desc(person)
```


Run str() function of file

```{r echo = TRUE }
str(person)
```


Run frequency of each variable

```{r echo = TRUE }
lapply(person, function(x) {

    if (is.numeric(x)) return(count(x))

    if (is.factor(x)) return(table(x))

})

```

Create categorical variables for analysis

```{r echo = TRUE }

person$sex <- factor(person$SEX)

person$notcov <- factor(person$NOTCOV)

person$delay <- factor(person$PDMED12M)

person$need <- factor(person$PNMED12M)

head(person$delay)
head(person$need)
```

Create new variables

```{r echo = TRUE }
person$need2 <- with(person, ifelse(person$PNMED12M == 1, 1, 0))

person$delay2 <- with(person, ifelse(person$PDMED12M == 1, 1, 0))

person$notcov2 <- with(person, ifelse(person$NOTCOV == 2, 1, 0))

person$need1 <- factor(person$need2)
person$delay1 <- factor(person$delay2)
person$notcov1 <- factor(person$notcov2)
```

Create new variables

```{r echo = TRUE }
person$medicaid <- with(person, ifelse(person$MEDICAID == 1 | person$MEDICAID == 2, 1, 0))
person$medicaid <- factor(person$medicaid)

person$medicare <- with(person, ifelse(person$MEDICARE == 1 | person$MEDICARE == 2, 1, 0))
person$medicare <- factor(person$medicare)

person$private <- with(person, ifelse(person$PRIVATE == 1 | person$PRIVATE == 2, 1, 0))
person$private <- factor(person$private)

person$schip <- with(person, ifelse(person$SCHIP == 1 | person$SCHIP == 2, 1, 0))
person$schip <- factor(person$schip)

person$ihs <- with(person, ifelse(person$IHS == 1 | person$IHS == 2, 1, 0))
person$ihs <- factor(person$ihs)

person$military <- with(person, ifelse(person$MILITARN == 1 | person$MILITARN == 2, 1, 0))
person$military <- factor(person$military)

person$public <- with(person, ifelse(person$OTHPUB == 1 | person$OTHPUB == 2, 1, 0))
person$public <- factor(person$public)

person$othgov <- with(person, ifelse(person$OTHGOV == 1 | person$OTHGOV == 2, 1, 0))
person$othgov <- factor(person$othgov)

person$single <- with(person, ifelse(person$SINGLE == 1 | person$SINGLE == 2, 1, 0))
person$single <- factor(person$single)

head(person$medicaid)
```


Create regression model for delayed care based on no insurance coverage

```{r echo = TRUE }
model <- glm(delay1~notcov1+sex+AGE_P, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(model)

```

Create regression model for needed care but did not receive care based on no insurance coverage

```{r echo = TRUE }
model2 <- glm(need1~notcov1+sex+AGE_P, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(model2)

```


Model  for types of insurance not seeking care

```{r echo = TRUE }
instypemodel <- glm(need1 ~ medicaid+medicare+private+schip+ihs+military+public+othgov+single, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(instypemodel)

```


Model  for medicare not seeking care

```{r echo = TRUE }
medicaremodel <- glm(need1 ~ medicare+sex+AGE_P, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(medicaremodel)

```



Model  for medicaid not seeking care

```{r echo = TRUE }
medicaidmodel <- glm(need1 ~ medicaid+sex+AGE_P, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(medicaidmodel)

```

Model  for private not seeking care

```{r echo = TRUE }
privatemodel <- glm(need1 ~ private+sex+AGE_P, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(privatemodel)

```

Model  for schip not seeking care

```{r echo = TRUE }
schipmodel <- glm(need1 ~ schip+sex+AGE_P, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(schipmodel)

```

Model  for ihs not seeking care

```{r echo = TRUE }
ihsmodel <- glm(need1 ~ ihs+sex+AGE_P, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(ihsmodel)

```

Model  for military not seeking care

```{r echo = TRUE }
militarymodel <- glm(need1 ~ military+sex+AGE_P, data = person, family = "binomial")

```
Output model results

```{r echo = TRUE }
summary(militarymodel)

```


Model  for public not seeking care

```{r echo = TRUE }
publicmodel <- glm(need1 ~ public+sex+AGE_P, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(publicmodel)

```

Model  for othgov not seeking care

```{r echo = TRUE }
othgovmodel <- glm(need1 ~ othgov+sex+AGE_P, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(othgovmodel)

```

Model  for single not seeking care

```{r echo = TRUE }
singlemodel <- glm(need1 ~ single+sex+AGE_P, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(singlemodel)

```


Histograms

```{r echo = TRUE }
# Histograms

hist(person$NOTCOV)
hist(person$PDMED12M)
hist(person$PNMED12M)
```

```{r echo = TRUE }
#Create dataset for testing correlation matrix

newdata <- person[,c('AGE_P','notcov2', 'SEX', 'delay2', 'need2')]

head(newdata)
```

Correlation matrix for insurance coverrage

```{r echo = TRUE }
#Correlation matrix 
cor(newdata)
```

Create new variables

```{r echo = TRUE }
person$sstypea <- with(person, ifelse(person$SSTYPEA == 1, 1, 0))
person$sstypea <- factor(person$sstypea)

person$sstypeb <- with(person, ifelse(person$SSTYPEB == 1, 1, 0))
person$sstypeb <- factor(person$sstypeb)


person$sstypec <- with(person, ifelse(person$SSTYPEC == 1, 1, 0))
person$sstypec <- factor(person$sstypec)

person$sstyped <- with(person, ifelse(person$SSTYPED == 1, 1, 0))
person$sstyped <- factor(person$sstyped)

person$sstypee <- with(person, ifelse(person$SSTYPEE == 1, 1, 0))
person$sstypee <- factor(person$sstypee)

person$sstypef <- with(person, ifelse(person$SSTYPEF == 1, 1, 0))
person$sstypef <- factor(person$sstypef)

person$sstypeg <- with(person, ifelse(person$SSTYPEG == 1, 1, 0))
person$sstypeg <- factor(person$sstypeg)

person$sstypeh <- with(person, ifelse(person$SSTYPEH == 1, 1, 0))
person$sstypeh <- factor(person$sstypeh)

person$sstypei <- with(person, ifelse(person$SSTYPEI == 1, 1, 0))
person$sstypei <- factor(person$sstypei)

person$sstypej <- with(person, ifelse(person$SSTYPEJ == 1, 1, 0))
person$sstypej <- factor(person$sstypej)


person$sstypek <- with(person, ifelse(person$SSTYPEK == 1, 1, 0))
person$sstypek <- factor(person$sstypek)

person$sstypel <- with(person, ifelse(person$SSTYPEL == 1, 1, 0))
person$sstypel <- factor(person$sstypel)


head(person$sstypel)
```

Histograms for coverage types

```{r echo = TRUE }
# Histograms

hist(person$SSTYPEA)
hist(person$SSTYPEB)
hist(person$SSTYPEC)
hist(person$SSTYPED)
hist(person$SSTYPEE)
hist(person$SSTYPEF)
hist(person$SSTYPEG)
hist(person$SSTYPEH)
hist(person$SSTYPEI)
hist(person$SSTYPEJ)
hist(person$SSTYPEK)
hist(person$SSTYPEL)

```




Model  for no cancer coverage and needed care

```{r echo = TRUE }
cancermodel <- glm(need1 ~ sstypec+sex+AGE_P, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(cancermodel)

```

Model  for no accident and needed care

```{r echo = TRUE }
accidentmodel <- glm(need1 ~ sstypea+sex+AGE_P, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(accidentmodel)

```

Model  for all coverages and needed care

```{r echo = TRUE }
plansmodel <- glm(need1 ~ sstypea+sstypeb+sstypec+sstyped+sstypee+sstypef+sstypeg+sstypeh+sstypei+sstypej+sstypek, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(plansmodel)

```


Create new variables

```{r echo = TRUE }
person$histop4 <- with(person, ifelse(person$HISTOP4 == 1, 1, 0))
person$histop4 <- factor(person$histop4)


head(person$histop4)
```

Model  for all Employer not offering coverage and needed care

```{r echo = TRUE }
empcovmodel <- glm(need1 ~ histop4 + sex + AGE_P, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(empcovmodel)

```

Model  for all Employer not offering coverage and delayed care

```{r echo = TRUE }
empcovmodel2 <- glm(delay1 ~ histop4 + sex + AGE_P, data = person, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(empcovmodel2)

```


Create Training and Testing Datasets

```{r echo = TRUE }
#create a list of random number ranging from 1 to number of rows from actual data and 70% of the data into training data  

data1 = sort(sample(nrow(person), nrow(person)*.7))

#creating training data set by selecting the output row values  
train<-person[data1,]

#creating test data set by not selecting the output row values
test<-person[-data1,]

```

Model cancer and needed care using training data

```{r echo = TRUE }
cancermodeltrain <- glm(need1 ~ sstypec+sex+AGE_P, data = train, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(cancermodeltrain)

```

Make predictions

```{r echo = TRUE }
#Make predictions
cancerpredict <- cancermodeltrain %>% predict(test, type = "terms")

head(cancerpredict)

summary(cancerpredict)
```




Model  for no accident and needed care using training set

```{r echo = TRUE }
accidentmodeltrain <- glm(need1 ~ sstypea+sex+AGE_P, data = train, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(accidentmodeltrain)

```

Make predictions

```{r echo = TRUE }
#Make predictions
accidentpredict <- accidentmodeltrain %>% predict(test, type = "terms")

head(accidentpredict)

summary(accidentpredict)
```






Model  for all coverages and needed care using training set

```{r echo = TRUE }
plansmodeltrain <- glm(need1 ~ sstypea+sstypeb+sstypec+sstyped+sstypee+sstypef+sstypeg+sstypeh+sstypei+sstypej+sstypek, data = train, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(plansmodeltrain)

```


Make predictions

```{r echo = TRUE }
#Make predictions
planspredict <- plansmodeltrain %>% predict(test, type = "terms")

head(planspredict)

summary(planspredict)
```





Model  for all Employer not offering coverage and needed care using training set

```{r echo = TRUE }
empcovmodeltrain <- glm(need1 ~ histop4 + sex + AGE_P, data = train, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(empcovmodeltrain)

```


Make predictions

```{r echo = TRUE }
#Make predictions
empcovpredict <- empcovmodeltrain %>% predict(test, type = "terms")

head(empcovpredict)

summary(empcovpredict)
```





Model  for all Employer not offering coverage and delayed care using training set

```{r echo = TRUE }
empcovmodeltrain2 <- glm(delay1 ~ histop4 + sex + AGE_P, data = train, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(empcovmodeltrain2)

```


Make predictions

```{r echo = TRUE }
#Make predictions
empcov2predict <- empcovmodeltrain2 %>% predict(test, type = "terms")

head(empcov2predict)

summary(empcov2predict)
```




